### 2023-04-17 paper.Rmd
continued

Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time if the underlying objectives are in conflict with each other [@ferligoj1992direct]. It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict. A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time. Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances. The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were simply set to 1), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.

>>>>

Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria at the same time if the underlying objectives are in conflict with each other [@ferligoj1992direct]. It is therefore unclear whether optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$ can yield a partitioning where both constituting criteria (mean and variance) are similar between groups, because it is unknown if these objectives tend to align with each other or are in conflict. A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time. Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances. The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were set to 1 when computing $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other.
